### 2015-03-13 learners.tex
Updates from ShareLaTeX

This section describes the learners used in this study. While there are many differences
in those learners, they have the same goal: to summarize a large data set into a small model\footnote{Technical note: proponents of instance-based reasoning might argue that
their goal is to retain, and not summarize, the training data.
However, 
even instance-based reasoners use some degree of summarizing when they learn
ways to bias the instance matching away from irrelevant training data.
Those bias tools include  generalizing points to ranges (to allow for partial matches)
or  clustering algorithm that restrict the case matching
to just the subset of old cases most similar  to new  cases.}.

For example,
consider some learner that
builds predictive rules from a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into $d=5$ divisions, then the 20 attributes of \fig{ck} can be combined into 
$2^{A*d} = 2^{5*20} > 10^{30}$
possible rules. The key thing to note here is that many of these rules will have similar effects
since, at least in software engineering, data sets containing rows $R$ and columns $C$ can contain much redundancy:
\bi
\item When $c \subset C$ columns of data that are correlated to each other, then those
$c$ columns can be replaced by one column.
\item When $r \subset R$ rows of data are very similar (e.g. they come from similar projects), 
then those $r$ rows can be replaced
by one cluster centroid that stands in $r$.
\item
Kocagenunil~\cite{me13a} and Papakroni~\cite{papa13} conduct  different experiments
to  show  that tables of software defect data in the form of \fig{ck} contain around 90\% redundant
data. Specifically, they show that such tables can be be pruned back to around
25\% of the columns and 10\% of the rows without losing the essential information in the data set.
\ei
Given this high level of redundancy, learners need only explore a small fraction of these $2^{A*d}$ rules.
Most learners exploit this fact to optimize their reasoning. Rather than explore all options,
they employ some {\em search bias} that lets them quickly focus on a small set of promising rules.

Different biases work best for different data sets so it is standard practice to allow an analyst
to control these biases via tuning parameters. For example, consider the search biases
$B_i$ of decision tree learner that :
\bi
\item Splits the data on some criteria $B_1$;
\item Assesses each split using some criteria $B_2$;
\item Selects which split(s) to explore further using $B_3$;
\item Recurses into each split until some stopping criteria $B_4$ is encountered.
\ei
These biases also appear when the learned decision tree makes a decision:
\bi
\item To declare a software module to be XXX
\ei

\section{Algorithm: Predictor and Tuner}

In order to conduct the experiment of this paper, we need one tool that can predict defects 
from the empirical data sets and a second tool to tune the built-in parameters associated with 
predictor. To compare the effects of the tuning process, we have three different predictors: 
WHERE-based Predictor, CART and Random Forest. Different Evolution(DE) as an optimizer 
is used as a tuner in this paper.

We choose CART and Random Forest as a predictor in this paper is motivated by 
\cite{lessmann2008benchmarking}, where the performance of both tools as predictors are 
significantly different based on authors' experiment as mentioned before. We'd like to 
investigate whether tuning can change such conclusion. WHERE-based Predictor is a new 
defect predictor based on WHERE\cite{menzies2013local} algorithm. A comparison with 
standard predictor like CART and Random Forest will better evaluate and judge the 
performance such new predictor. As for the tuner, there're many heuristic optimization 
algorithms in wild. However, DE is a good but maybe not the best candidate for the tuning 
process considering different performance measurements. To determine which optimizer is 
fitable and results in better performance is beyond this work scope. We leave it to future 
works. The rest of this section will describe each tool applied in this work.

 \subsection{WHERE-based Learner}
\textbf{WHERE-based Learner} is composed of WHERE clustering algorithm and CART 
decision tree algorithm. The key idea of WHERE-based learner is that  instead of training the 
CART decision tree based on the class labels associated with each training sample, it's using 
CART to build decision trees based on the cluster labels, which are generated by the WHERE 
clustering algorithm.

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{r@{~}|r@{~}rr|r@{~}rr|r@{~}rr}
    & \multicolumn{3}{c|}{WHERE}&\multicolumn{3}{c|}{CART}& \multicolumn{3}{c}{Random Forest}\\\cline{2-10} 
    Goal & tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive&$\frac{\mathit{tuned}}{\mathit{naive}}$\\\hline
pd&71.4&1.2&60&4.4&0.1&44&7.1&0.2&36\\ 
prec&82.8&1.4&59&3.9&0.1&39&7.4&0.2&37\\
f&86.1&1.2&72&3.3&0.1&33&6.7&0.2&34
  \end{tabular}
  \caption{For one train/test set pair (ANT),  time (in seconds) spent tuning,
  for different goals, or just running the ``off-the-shelf'' learner.}
\end{figure}

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
\begin{tabular}{r|rrr}
goal &	WHERE	&CART&	Random Forest\\\hline
f	&15	&13	&9\\ 
pd	&15	&16	&8\\ 
prec&	14	&14	&8
\end{tabular}
\caption{How much slower is tuning vs just running the default
settings of the learner? E.g. bottom right, tuning to increase
precision with Random Forests is eight times slower than
just running Random Forests.}
\end{figure}

% Given N instances, instead of building decision tree directly with CART, we use WHERE to 
%cluster instances. 
WHERE is a fast clustering algorithm designed by {\it menzies} for finding software artifacts 
with similar attributes. It clusters data on dimensions synthesized along the axis of greatest 
variability in the data. The way WHERE used to find such dimension is a linear-time heuristic 
called ``FASTMAP" proposed by Faloutsos \& Lin\cite{faloutsos1995fastmap}. ``FASTMAP" 
randomly picks one instance $Z$; find the instance {\it east} $X$ that's furthest away form $Z$; find 
the instance {\it west} $Y$ that's furthest away from {\it east} $X$. Next, project all the remaining 
points onto the line drawn between $X$ and $Y$.  The line $\overline{XY}$ is an approximation of the 
first component found by PCA. As shown in Figure \ref{Where}, $X$ and $Y$ are the furthest points found by ``FASTMAP'' and $\overline{XY}$ is of length $c$. Each point in this figure has a distance $a$
to $X$ and distance $b$ to $Y$. According to the coisne rule and Pythagoras, each instance will be mapped 
into 2-dimension by the following equations.
\begin{equation}
\begin{split}
x  &= (a^2 +c^2-b^2)/(2c)\\ 
   y  &= \sqrt{a^2 -x^2}
\end{split}
\end{equation}

Then choose the median point $\hat{x}$ as the split and 
recursively divide all the instances into {\it west} and {\it east} clusters until the number of 
instances within each cluster is less than specific minimum size. The representation of the 
clustering result is a tree, we call it Where-clustering tree. Such tree will be pruned if applicable 
in some cases. Finally, cluster labels  will be assigned to instances in each of those 
clusters(leaves).

\begin{figure}[!ht]
\includegraphics[scale=0.5]{where.png}
\caption{WHERE algoirthm }
\label{Where}
\end{figure}
Then based on the attributes and the cluster labels in each instances,  we build a 
decision tree based on CART algorithm. In each leaf of the tree, there're several instances 
falling in. This will be the model trained for defect prediction. During testing process, a new 
instance comes in and traverses the tree according to 
values of the its attributes. If the instance could reach one leaf of the tree, the predicted value 
of this instance will be the mean  of all the training data values associated with this leaf. 
Otherwise if stops at intermediate nodes of tree, the predicted value will be the mean of the all 
the training data value below this node. After estimating all the number of defectives in the test 
data, whether each instance contains defectives will be determined by comparing with a threshold value. If the predicted value is greater than threshold, the Where-based learner will predict this instance as defective otherwise non-defective.

%\begin{algorithm}
%	\caption{Pesudocode for WHERE Algorithm }
%	\label{alg:WHERE}
%	\begin{algorithmic}[1]
%	\Require $data$
%	\Ensure $tree$
%	\Function {Fastmap}{$ data$}
%	\State $one \gets$ any$(data)$
%	\State $west \gets$ furthest$(data, one)$
%	\State $east \gets$ furthest$(data, west)$
%	\State $c \gets$ distance($west, east$)
%	\EndFunction
%	\end{algorithmic}            
%\end{algorithm}

\subsection{CART}
\textbf{CART} is an {\em iterative dichotomization} algorithm
that finds the attribute that most divides the data such that
the variance of the goal variable in each division is
minimized\cite{breiman84}.
The algorithm then recurses on each division.
Finally, the cost data in the leaf divisions
is averaged to generate the estimate.

\subsection{Random Forests}

Breiman's website describes \textbf{Random Forests} as
follows\cite{brieman00}. "Random Forests grows many classification
trees. To classify a new object from an input vector, put the input
vector down each of the trees in the forest. Each tree gives a
classification, and we say the tree "votes" for that class. The forest
chooses the classification having the most votes (over all the trees
in the forest). Each tree is grown as follows.
If the number of cases in the training set is N, sample N cases at
random - but with replacement, from the original data. This sample
will be the training set for growing the tree.
Also, if there are M input variables, a number m$<<$M is specified such
that at each node, m variables are selected at random out of the M and
the best split on these m is used to split the node. The value of m is
held constant during the forest growing.
Finally, each tree is grown to the largest extent possible (there is
no pruning)."

%Suppose the cordinates of X and Y are $(0,0)$ and $(0,c)$, which are east and west 
%instances within N samples. From the Pythegoras and cosine rule, each instance is projected 
%at the point(x,y)
%
%\begin{equation}
%	\begin{split}
%	x &= (a^2 + c^2 - b^2)/(2c) \\
%	y &= \sqrt{a^2 - x^2} 
%	\end{split}
%\end{equation}