### 2015-03-15 dm101.tex
Updates from ShareLaTeX

Human programmers are clever, but flawed. Coding  adds functionality, but also defects.
Hence, software sometimes crashes (perhaps at the most awkward or dangerous moment) or delivers
the wrong functionality. For a very long list of software-related errors,
see  Peter Neumann's ``Risk Digest'' at catless.ncl.ac.uk/Risks.

Since programming inherently
introduces defects into  programs, is important to thoroughly {\em test} software before it is {\em used}.
Such testing  can be very expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
Lowry et al. warn that  
the state space explosion problem imposes
strict limits on how much a system can be explored
via automatic formal methods~\cite{lowrey98}.
For standard black-box testing methods,
a {\em linear} increase
in the confidence $C$ that we have found all defects
can take {\em exponentially} more effort.
For example, for one-in-a-thousand detects,
moving $C$ from  
90\% to 98\% nearly doubles the number of  tests (2301 to   3910 black box
probes, respectively)\footnote{A randomly selected 
input to a program will find a fault with probability $p$.
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=log(1 -
C)/log(1-p)$. For example, $N(0.90,10^{-3})=2301$.}.

Exponential costs quickly exhaust finite resources.
Standard practice is to apply the best
available assessment methods on the sections of the program that the
best available domain knowledge declares is most critical.  We endorse
this approach.  Clearly, the most critical sections require the best
known assessment methods. However, this focus on certain sections
can blind us to defects in other areas.
Therefore, standard practice should be augmented
with a  {\em
lightweight sampling policy} to explore the rest of the system.  This
sampling policy will always be incomplete.
Nevertheless, it is the recommended option when
resources do not permit a complete assessment of the whole system.

One such lightweight sampling policy is defect predictors learned from static code attributes.
Given software described using (for example) the attributes of \fig{ck}, it is possible
to use standard data mining technology to learn where the probability of software defects is highest.
Such defect  defect predictors learned from
static code attributes are   {\em easy to
use}, {\em widely-used}, and {\em useful} to use.

{\em Easy to use:} Static code attributes can be automatically collected, even for very large systems~\cite{nagappan05}.
Other methods, like  manual code reviews, are far slower and far more labor-intensive.
For example, depending on the review methods 8 to 20 LOC/minute can be
inspected and this effort repeats for all members of the review team,
which can be as large as four or six~\cite{me02f}.

%%%%%%%%%%%%%%%% list of parameters%%%%%%%%%%%%%%%%%%%%%
\renewcommand\arraystretch{1.2}
\begin{figure*}[t!]
\scriptsize
  \centering
	\begin{tabular}{|c|c|c|c|l|}
	\cline{1-5}
	\begin{tabular}[c]{@{}c@{}}Learner \\ Name\end{tabular} & Parameters & Default &\begin{tabular}[c]{@{}c@{}}Tuning\\ Range\end{tabular}& 
\multicolumn{1}{c|}{Description} \\ \hline
	\multirow{8}{*}{\begin{tabular}[c]{@{}c@{}}Where-based\\ Learner\end{tabular}} 
	& threshold & 0.5 &[0.01,1]& The value to determine defective or not .\\ \cline{2-5} 
	& infoPrune & 0.33 &[0.01,1]& The percentage of features to consider  for the best 
split to build CART tree\footnote{Since the Where-based learner will build two trees, the first 
one is for clustering and the second one is building prediction model. we explicitly call Where-
clustering tree and CART tree, respectively}. \\ \cline{2-5} 
	 & min\_sample\_split & 4& [1,10]& The minimum number of samples required to split an internal node of
CART tree. \\ \cline{2-5} 
	 & min\_Size & 0.5 &[0.01,1]& \begin{tabular}[c]{@{}l@{}}Finds min\_sample 
in Where-clustering trees  using  ${n\_samples}^ {min\_Size}$.
\end{tabular} \\ \cline{2-5} 
    & wriggle & 0.2 &[0.01, 1] & The threshold to determine which branch in  Where tree to be pruned\\ \cline{2-5}
	 & depthMin & 2 & [1,6]&The minimum depth of the tree below which no pruning for Where-
clustering tree. \\ \cline{2-5} 
	 & depthMax & 10 &[1,20]& The maximum depth of the Where-clustering tree. \\ \cline{2-5} 
	 & wherePrune & False &T/F& Whether or not to prune the Where-clustering tree. \\ \cline{2-5}
	 & treePrune & True &T/F& Whether or not to prune the classification tree built by CART. \\ \cline{2-5} 
\hline
\multirow{4}{*}{CART} & threshold & 0.5 &[0,1]& The value to determine defective or not. \\ \cline{2-5} 
	 & max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5}

	 & min\_sample\_split & 2 &[2,20]& The minimum number of samples required to split an 
internal node. \\ \cline{2-5} 
	 & min\_samples\_leaf & 1 & [1,20]&The minimum number of samples required to be at a leaf 
node. \\ \cline{1-5}  
       \multirow{5}{*}{\begin{tabular}[c]{@{}c@{}}Random \\ Forests\end{tabular}}  & threshold & 0.5 & [0.01,1] & The value to determine defective or not. \\ 
\cline{2-5} 
	 & max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5} 
	 & max\_leaf\_nodes & None &[1,50]& Grow trees with max\_leaf\_nodes in best-first fashion. \\ \cline{2-5} 
	 & min\_sample\_split & 2 &[2,20]& The minimum number of samples required to split an 
internal node. \\ \cline{2-5} 
	 & min\_samples\_leaf & 1 &[1,20]&The minimum number of samples required to be at a leaf 
node. \\ \cline{2-5} 
	 &  n\_estimators & 100 & [50,150]&The number of trees in the forest.\\ \cline{2-5}
	 \hline

	\end{tabular}
    \caption {List of parameters to be tuned.}
\label{fig:parameters}
\end{figure*}
{\em Widely used:} Many researchers and industrial practitioners  use static attributes to guide software 
quality predictions.
 Defect prediction models have been reported
to have been used at Google~\cite{lewis13}.
Verification and validation (V\&V) textbooks
(\cite{rakitin01}) advise using static code complexity attributes
to decide which modules are worthy of manual inspections.  
For several  years, one of us (Menzies) worked on-site at the NASA software Independent Verification
and Validation facility
and he
knows of several large government software contractors that won't
review software modules {\em unless} tools like McCabe predict that
they are fault prone.

{\em Useful:}
Defect predictors often  find the location of  70\% (or more)
of the defects in code~\cite{me07b}.
Defect predictors developed at NASA~\cite{me07b} have also been used in software development companies outside the US (in Turkey). When the inspection teams focused on the modules that trigger the defect predictors, they found up to 70\% of the defects using just 40\% of their QA effort (measured in staff hours)~\cite{tosun10}.
A subsequent study on the Turkish software
compared how much code needs to be inspected using
random selection vs. selection via defect
predictors. Using random testing, 87\% of the files
would have to be inspected in order to detect 87\%
of the defects. Further, if the inspection process
was restricted to the 25\% of the files that trigger
the defect predictors, then 88\% of the defects
could be found. That is, the same level of defect
detection (after inspection) can be achieved using
$(87-25)/87=71$\%less effort
\cite{tosun09}.

Defect predicting technology has been
commercialized in many tools including {\it Predictive}~\cite{turner06}, a
 product suite to analyze and predict
defects in software projects. Predictive was observed to
highlight similar issues to those found   with the more expensive tools. Significantly,
Predictive was able to faster process a larger code
base than the more expensive tool~\cite{turner06}.

The success of this method in  predictors in finding bugs is   markedly
higher than other currently-used
industrial
methods such as manual code reviews. For example, 
a  panel at {\em IEEE Metrics
2002}~\cite{shu02} concluded that manual software  reviews can find ${\approx}60\%$ 
of defects.
In other work, 
Raffo documents the typical    defect detection capability of
industrial review methods:   around 50\%
 for full Fagan inspections~\cite{fagan76} to
21\% for less-structured inspections.

Note only do static code defect predictors perform well compared to manual methods,
they also are competititve with certain automatic methods.
At ICSE'14, Rahman et al.~\cite{rahman14:icse} compared:
\bi
\item The static code analysis tools FindBugs, Jlint, and Pmd;
\item Against static code defect predictors
(which they called ``statistical defect prediction'') build using logistic regression.
\ei
They found  no significant differences in the cost-effectiveness
of these  approaches. Given this equivalence, it is significant to note that 
static code defect prediction can be quickly adapted to new languages by building lightweight
parsers that find   information like \fig{ck}. The same is not true for   static code analyzers-- these need  extensive modification before they can be used on new
languages.